A Comprehensive Dataset of U.S. Laws

### 2023-08-10 draft/v1.tex

In the US federal system, laws in the sense of ordinary usage can originate from many sources, including Courts, the President, agencies, and so forth. Yet when scholars speak more formally about the set of "US laws," they typically intend to refer to the set of authoritative documents enacted via the legislative process by the two chambers of the US Congress, with the possible input of the President. When Congress acts legislatively, it may do so by issuing a variety of documents that go by different names, have different formal features, and follow different processes. At present writing, the set of allowable actions include bills, joint resolutions, concurrent resolutions, and simple resolutions. Of these, only bills and joint resolutions have the general and binding character that one typically associates with laws. And even then, there are exceptions within these categories. "Private bills" granting relief in one-off cases are typically not what scholars have in mind when one refers to ``laws'', nor are joint resolutions approving constitutional amendments which are not ratified by the states (e.g. the Equal Rights Amendment) really laws. That said, the set of enacted bills and joint resolutions is a relatively precise definition of what scholars \textit{should} mean when they talk about the set of US federal laws.

Studies of lawmaking in US political science are not new. Indeed, a very important debate in political science over the past several decades has to do with whether unified government, where the President and Congress are controlled by the same party, is better or worse for legislative productivity than divided government, where these organs are not exclusively in controlled by the same political party. This debate was initiated by Mayhew, who published a list of ``major'' legislative enactments identified by the set of laws that were mentioned in the editorial page of the New York Times and other prominent newspapers. This list of major enactments has been updated nearly annually since its original publication in DATE. Of course, the coverage of major legislative enactments is obviously limited as a database of all federal laws, both because it does not cover laws that fail to draw the attention of editorial pages, nor does it cover laws enacted before the existence of outlets such as the New York Times.  Indeed, PERIOD COVERED.

Addressing these weaknesses has been the subject of at least two other research projects. The Policy Agendas Project does X. Their focus on covering the subject matter of lawmaking is an important undertaking, although arduous, and it is unsurprising given the difficulties of this undertaking that they have only managed to cover X-Y. Ansolabhere Palmer Schneer devise novel methods for extending the list of major enactments backward in time. DESCRIBE HOW THEY DO IT. Still, their coverage is hardly exhaustive.

There are several reasons why it is desirable to have a more complete dataset of US lawmaking than has previously existed. In particular, minor legislation is usually major to somebody, otherwise they would not go through the trouble of passing it. While focusing on the big pieces of legislation is perhaps appropriate for thinking about how party control effects the ability of collective government to solve the big problems, a more granular perspective is really necessary to think about how special interests may whittle away the gains made by broad coalitions through smaller pieces of legislation, or perhaps how policy fails to make the necessary incremental adjustments to keep with the times. The lack of a unified and complete dataset also frustrates progress of the discipline, as these authors do not necessarily use the same system to refer to particular laws, making US law datasets used by social scientist less interoperable than they should be.

Our project seeks to build a comprehensive and easily updated list of US laws and their revision histories based on publicly available sources. The two key sources we leverage are the Statutes at Large, for laws existing before 2012, and Congress.gov, for more recent laws. The Congress.gov component is downloadable by webscraping and is relatively uncomplicated.  The more challenging part of the project involves cataloguing the older set of laws contained within the Statutes at Large. Surprisingly, the US government does not make electronic copies of the Statutes at Large available prior to 1951,\footnote{https://www.govinfo.gov/app/collection/statute/2016}, despite the fact that there were 64 earlier volumes containing laws that in many cases have not been amended, for example the law from the 1870s that formed the basis of an indicment against Donald Trump for ``conspiracy against rights.'' Despite the surprising limited availability of these foundational documents, a database widely used by legal scholars, HeinOnline, has digitized all these books in their entirety and provided meta data on their table of contents. Laws are identified in the statutes of laws in various and not perfectly ways across the decades. For this reason, we supply a more consistent set of identifiers that, going forward, allows laws to be easily matched by researchers. Additionally, we include a set of text data that corresponds to these laws, although particularly for earlier laws these will be collected via OCR.

@@

@@

\section*{Usage Notes}

@@

\bibliography{sample}

>>>>

In the US federal system, laws in the sense of ordinary usage can originate from many sources, including Courts, the President, agencies, and so forth. Yet when scholars speak more formally about the set of "US laws," they typically intend to refer to the set of authoritative documents enacted by the two chambers of the US Congress by following the legislative process. When Congress acts legislatively, it may do so by issuing a variety of documents that go by different names, have different formal features, and follow different processes. At present writing, the set of allowable actions include bills, joint resolutions, concurrent resolutions, and simple resolutions. Of these, only bills and joint resolutions have the general and binding character that one typically associates with laws. And even then, there are exceptions within these categories. ``Private bills" granting relief in one-off cases\footnote{For example, last year Congress enacted a private law allowing Alaska-woman Rebecca Trimble to obtain permanent resident status, despite the fact the paperwork her adoptive parents had used to bring her to the United States from Mexico had been defective.} are typically not what scholars have in mind when one refers to ``laws'', nor are joint resolutions approving constitutional amendments which are not ratified by the states (e.g. the Equal Rights Amendment) really laws. That said, the set of enacted bills and joint resolutions is a relatively precise definition of what scholars \textit{should} mean when they talk about the set of US federal laws.

Studies of lawmaking in US political science are not new. Indeed, a very important debate in political science over the past several decades has to do with whether unified government, where the President and Congress are controlled by the same party, is better or worse for legislative productivity than divided government, where these organs are not exclusively in controlled by the same political party. This debate was initiated by Mayhew (1991) \nocite{Mayhew1991}, who published a list of ``major'' legislative enactments identified by the set of laws that were mentioned in the editorial page of the New York Times and other prominent newspapers. This list of major enactments has been updated annually since its original publication in 1991 and is, at present writing, current from 1947 through 2022. Of course, the coverage of major legislative enactments is obviously limited as a database of all federal laws, both because it does not cover laws that fail to draw the attention of editorial pages, nor does it cover laws enacted during the New Deal or earlier.

Addressing these weaknesses has been the subject of at least two other major research projects. The Policy Agendas Project is a long-standing collaborative research effort aimed at building multiple datasets for tracking and comparing which issues manage to attract the attention of government across. It covers many national and subnatnional systems, including hte United States federal government. In particular, their database of bills covers "more than 400,000 bills introduced by the US Congress." Importantly, they also have coded each bill according to the subject-matter coding system of the larger Policy Agenda Project. Their focus on covering the subject matter of lawmaking is an important undertaking, although arduous, and it is unsurprising given the difficulties of this undertaking that they have only managed to cover the period 1947-2022. The second project with relevant and similar prior data collection is due to Ansolabehere, Palmer, and Schneer.\cite{ansolabehere_palmer_schneer_2016,ansolabehere2018divided} Their data collection, which was conducted under the auspices of an undergraduate course at Harvard called ``"What Has Congress Done," crowd-sourced the task of identifying significant legislation to students, each of whom was responsible for producing a list of significant legislation enacted under each of 22 potentially assigned decades. The researchers provided a list of secondary sources, as well as guidance and quality control efforts. Similar to Mayhew, their criteria for inclusion were (a) was the law considered significant at the time of enactment, and (b) does the law seem significant in historical perspective. Because of their differing methods, the two sources do not completely overlap in the 1947-2022 period. Although the temporal coverage for the Ansolabehere, Palmer, and Schneer dataset is very comprehensive, it contains a small percentage of all laws enacted by the US Congress.

% A few OTHER PAPERS
% Clinton and Lapinski (2006)
% Stathis (2014)
% Howell, William, Scott Adler, Charles Cameron, and Charles Riemann (2000) “Divided government
and the legislative productivity of Congress, %1945–94.” Legislative Studies Quarterly 25 (2): 285–
%312

There are several reasons why it is desirable to have a more complete dataset of US lawmaking than has previously existed. In particular, minor legislation is usually major to somebody, otherwise they would not go through the trouble of passing it. While focusing on the big pieces of legislation is perhaps appropriate for thinking about how party control effects the ability of collective government to solve the big problems, a more granular perspective is really necessary to think about how special interests may whittle away the gains made by broad coalitions through smaller pieces of legislation, or perhaps how policy fails to make the necessary incremental adjustments to keep with the times. There is notable inconsistency in what laws are included as significant using all these methods. The lack of a unified and complete dataset also frustrates progress of the discipline, as these authors do not necessarily use the same system to refer to particular laws, making US law datasets used by social scientist less inter-operable than they should be.

Our project seeks to build a comprehensive and easily updated list of US laws and their revision histories based on publicly available sources. The two key sources we leverage are the Statutes at Large, for laws existing before 2012, and Congress.gov, for more recent laws. The Congress.gov component is downloadable by webscraping and is relatively uncomplicated.  The more challenging part of the project involves cataloguing the older set of laws contained within the Statutes at Large. Surprisingly, the US government does not make electronic copies of the Statutes at Large available prior to 1951,\footnote{https://www.govinfo.gov/app/collection/statute/2016}, despite the fact that there were 64 earlier volumes containing laws that in many cases have not been amended. Despite the surprising limited availability of these foundational documents, a database widely used by legal scholars, HeinOnline, has digitized all these books in their entirety and provided meta data on their table of contents. Laws are identified in the statutes of laws in various and not perfectly ways across the decades. For this reason, we supply a more consistent set of identifiers that, going forward, allows laws to be easily matched by researchers. We describe these identifiers in more detail in the usage section below. Additionally, we include a set of text data that corresponds to these laws, although particularly for earlier laws these are collected via OCR.

@@

We rely on two primary sources for our comprehensive database of US laws: the U.S. Statutes at Large, as disseminated via HeinOnline, and Congress.gov.

For our data collection based on HeinOnline, we began by scraping the table of contents of each volume of the statues at large, excepting volumes 6-8 which contain a collection of early treaties and private laws. Examination of these tables reveals that laws are identified using various citation systems depending on era. In particular,

\begin{itemize}
\item The currently practiced identification system begins on Jan 7, 1959 with the 86th Congress. In this system, laws are identified by the number of the Congress and the law number within Congress. An example is ``Public Law 105-89, An Act: to promote the adoption of children in foster care.'' 
\item Initially, laws were identified by a different system. From the 1st through the 56th Congress, which concluded March 3, 1901, laws were identified in the Statutes at Large by chapter as well as Congress and session (e.g. ``Chapter 15, 56 Congress, Session 1, An Act: Relating to Cuban vessels.") 
\item For the 57th through 85th congress, the Statutes at Large inconsistently use both systems, with the important caveat that chapters and public laws are sometimes recycled over the two sessions of a particular Congress.\footnote{Or worse, the same session. Indeed, PL 65-246 refers to both a law ``Providing for the transportation from the District of Columbia of governmental employees whose services no longer are required." and also a law intended ``To authorize the sale of certain lands to school district numbered twenty-eight, of Missoula County, Montana."} Careful attention to chapters, public laws, and sessions does disambiguate all laws, however these details are very particular and subtle. 
\end{itemize}

Following these citation practices, we used a rule-based parser to extract key pieces of information to identify the law. Technical validation efforts revealed that there were occasional errors of transcription in our source data, for example mis-identification of chapters associated with laws or including page numbers that were not correct. Those errors that were found in the source data we corrected to conform with the text image captured on the HeinOnline site.

@@

We offer two primary approaches to support the technical validation of our dataset. The first is to compare the total counts of laws by Congress against previously published totals while the second validation exercise is to show that all of the laws described in five distinct sources of major laws are, indeed, present in our dataset.

The reference counts for total laws that we rely on comes from Appendix F reported in Galloway and Wise's History of the House of Representatives, and also previously used for validation and analysis by Ansolabhere, Palmer, and Schneer.\cite{ansolabehere_palmer_schneer_2016} Our definition of law is slightly more capacious than theirs, in the sense that we also consider "Joint Resolutions" laws, and the authors admit their own counts combine acts and resolutions (the latter of which are not, in our view, best understood as laws) from the 77th Congress and after. Figure \ref{fig:totals} compares the counts for this key subset of our data. Note that the total deviation in counts is X. An undergraduate research assistant further investigated the difference in counts of laws in Congress A and B. In both cases, the number of laws that the research assistant counted in that Congress was the same as the number we provided.

A second 
\

\section*{Usage Notes}  \label{sec:usage}

@@

\bibliography{biblio}
